The Regional Effects of Marine Protected Areas

### 2019-10-09 documents/ovando-regional-effects-of-mpas.tex
Initial commit of cleaned up regional effects repo

---
title: The Regional Effects of Marine Protected Areas
author:
  - name: Daniel Ovando
    affiliation: a,1
    email: danovan@uw.edu
  - name: Jennifer E. Caselle
    affiliation: b
  - name: Christopher Costello
    affiliation: b   
  - name: Olivier Deschenes
    affiliation: b  
  - name: Steven D. Gaines
    affiliation: b
  - name: Ray Hilborn
    affiliation: a
  - name: Owen Liu
    affiliation: b        
address:
  - code: a
    address: University of Washington, School of Aquatic and Fishery Sciences
  - code: b
    address: University of California, Santa Barbara
corresponding_author:
  - code: 1
    text: "To whom correspondence should be addressed. E-mail: danovan@uw.edu"
lead_author_surname: Ovando
author_contributions: |
  D.O., S.G., and R.H. developed structure of simulation model. D.O., O.D., C.C., and J.C. developed estimation strategy. J.C. provided support on collecting and interpreting data. All simulations, statistics, and sensitivity analyses performed by D.O. and O.L. . All authors contributed to the manuscript. 
conflict_of_interest: |
  R.H's research program receives funding from environmental NGOs, foundations, fishing industry, governments and international agencies. All of these can be interpreted as a conflict of interest when evaluating fisheries policy. C.C., S.G., and O.D's research program includes funding from environmental NGOs and foundations with an interest in ocean conservation and food security
abstract: |
  Marine Protected Areas (MPAs) cover between 3-7% of the world's oceans, up from less than 1% in the year 2000. The Convention on Biological Diversity calls for 10% of coastal waters to be protected inside MPAs by 2020, while the International Union for Conservation of Nature calls for 30% protection by 2030. It is often clear that MPAs produce conservation benefits inside their borders, but many MPAs are also justified on the grounds that they also benefit the broader region outside their borders. Despite the expanding role of MPAs in marine resource management, we lack a clear understanding of what the regional effects of MPAs might be, and how we might detect them in scientific studies. We used a bio-economic simulation model to demonstrate how the regional effects of MPAs are affected by a suite of economic and environmental drivers and show how commonly used methods to measure inside-MPA effects are likely to fail to capture the regional impacts of protection. We present an empirical strategy for estimating regional MPA effects on biomass density using 16 years of data from inside and outside of MPAs within the Channel Islands, California, USA. We find no statistically clear effect of MPAs on aggregate mean biomass densities of targeted finfish throughout the study region. To make meaningful progress on MPA design and evaluation, the next generation of MPA science must address the challenge of understanding and estimating the regional effects of MPAs across the diversity of species in the ecosystem.
significance: |
  Healthy marine ecosystems are critical to the well-being of the planet. Marine protected areas, parts of the oceans protected from human activities such as fishing, are increasingly being used in an effort to conserve and manage these ecosystems. Our study pairs theory with empirical methods to examine the regional effects of MPAs on conservation outcomes. Using simulations, we show that the effects of marine protected areas can be highly variable and dependent on human drivers. Building off of MPA theory, we find no statistically clear effect of a network of MPAs, which our simulation analysis suggests is entirely reasonable. Our results lay a foundation for future research on the design and monitoring of marine protected areas. 
acknowledgements: |
  This study would not be possible without the work provided by PISCO divers over the years. DO thanks Cody Szuwalski, Julia Lawson, and André Punt for helpful comments and technical support. 
keywords:
  - Marine Protected Areas
  - Conservation
  - Bio-economic modeling
  - Program Evaluation
## must be one of: pnasresearcharticle (usual two-column layout), pnasmathematics (one column layout), or pnasinvited (invited submissions only)
pnas_type: pnasresearcharticle
bibliography: My Library.bib
csl: pnas.csl
lineno: true
header-includes:
   - \usepackage{eso-pic,graphicx,transparent}
output:
  rticles::pnas_article: default
  bookdown::word_document2: default
  bookdown::pdf_document2: default
  bookdown::html_document2: default
params:
  run_name: ["v5.1"]
  sim_years: [50]
  burn_years: [20]
  num_patches: [25]
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dev = "CairoPDF", dpi = 600,
                      cache = FALSE)
# knitr::opts_knit$set(root.dir = "/Users/danovando/PhD/zissou") # not entirely reprodcible hack to deal with knit child in dissertation.. not idealk

```

<!-- why is it going to be hard ot detect regional effects of MPAs few if any empirical studies at net regional effects. does this mean that i doesn't exist, or impossible tot est, why? three fundamental stumbling blocks -->

```{r, include=FALSE}
library(sf)
library(ggmap)
library(viridis)
library(hrbrthemes)
library(scales)
library(patchwork)
library(rpart)
library(extrafont)
library(caret)
library(ggsci)
library(rEDM)
library(tidyverse)
extrafont::loadfonts()
functions <- list.files(here::here("functions"))

walk(functions, ~ here::here("functions", .x) %>% source()) # load local functions
save_plots <- TRUE

# sim_years <- 50
# 
# burn_years <- 20
# 
# num_patches <- 25

# run_name <- "v5.0"

run_name <- params$run_name

sim_years <- params$sim_years

burn_years <- params$burn_years

num_patches <- params$num_patches

run_dir <- here::here("results", run_name)

experiment_dir <- here::here("results",run_name, "experiments")

load(file = here::here("results",run_name, "rawish_zissou_data.Rdata"))

load(file = here::here("results",run_name, "model_runs.Rdata"))

processed_grid <-  read_rds(file.path(run_dir,"filtered_processed_grid.rds"))

load(file = file.path(run_dir ,"sim_grid.Rdata"))

cdfw_catches <-
  read_csv(file = here::here("data", 'cdfw-catches.csv')) %>%
  group_by(sci_name, year) %>%
  summarise(catch = sum(pounds_caught, na.rm = T)) %>%
  left_join(
    life_history_data %>% select(taxa, classcode, commonname) %>% mutate(taxa = tolower(taxa)),
    by = c("sci_name" = "taxa")
  ) %>%
  filter(!is.na(classcode))

# cdfw_catches %>%
#   group_by(commonname, classcode, sci_name) %>%
#   summarise(catch =sum(catch, na.rm = TRUE)) %>%
#   mutate(in_analysis = classcode %in% unique(cip_data$classcode)) %>%
#   View()

density_ratios <- readRDS(file.path(run_dir,"density_ratios.rds"))

outcomes <- readRDS(file.path(run_dir,"outcomes.rds"))

outcomes <- outcomes %>% 
  filter(!is.na(mpa_effect))

density_ratios <- density_ratios %>% 
  filter(!is.na(biased_density_ratio))

   eqo <- outcomes %>%
     filter(year == max(year))

facet_labels <- c(
  mpa_size = "Range in MPA",
  depletion = "Depletion"
)

channel_islands <- readRDS(here::here("data","channel_islands_map.RDS"))

ca_mpas <- sf::st_read(here::here("data","MPA_CA_Existing_160301")) %>%
  rmapshaper::ms_simplify() %>%
  sf::st_transform(crs = 4326)

zissou_theme <-
  theme_ipsum(
    base_size = 12,
    axis_title_size = 12,
    strip_text_size = 12,
    base_family = "Fira Sans"
  )

testplot <- ggplot()

panel_height = unit(1,"npc") - sum(ggplotGrob(testplot)[["heights"]][-3]) - unit(4,"line")

gc <- guide_colorbar(frame.colour = "black",
                     ticks.colour = "white",
                     barheight = panel_height)

  hgc <- guide_colorbar(frame.colour = "black",
                       ticks.colour = "black",
                       barwidth = 10)

    wide_hgc <- guide_colorbar(frame.colour = "black",
                       ticks.colour = "black",
                       barwidth = 25)

theme_set(zissou_theme)

models_worked <- model_runs$tmb_fit %>% map("error") %>% map_lgl(is_null)

model_runs <- model_runs %>%
  filter(models_worked) %>%
  mutate(tmb_fit = map(tmb_fit,"result")) %>%
  mutate(processed_fits = map(tmb_fit, process_fits)) %>%
  mutate(did_plot = map(processed_fits, "did_plot"))

base_run <- model_runs %>%
filter(var_names == "pisco_a", mpa_only == FALSE, center_scale == TRUE)

mpa_run <- model_runs %>%
  filter(data_source == "pisco",
         var_names == "pisco_a",
         mpa_only == TRUE)

kfm_run <- model_runs %>%
  filter(data_source == "kfm")

fitted_data <- base_run$data[[1]]

zissou_fit <- base_run$tmb_fit[[1]]

report <- base_run$tmb_fit[[1]]

site_data <- read_csv(here::here("data",'Final_Site_Table_UCSB.csv')) %>%
  magrittr::set_colnames(., tolower(colnames(.))) %>%
  select(
    site,
    side,
    mpagroup,
    mpa_status,
    reserve,
    region,
    year_mpa,
    mpaareanm2,
    lat_wgs84,
    lon_wgs84
  ) %>%
  rename(lat_wgs84 = lon_wgs84,
         lon_wgs84 = lat_wgs84) %>%
  unique() %>%
  mutate(eventual_mpa = (year_mpa > 0))

seen_non_nested_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_non_nested_betas")) %>%
  rename(group = variable) %>%
  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())

seeing_non_nested_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_non_nested_betas")) %>%
  rename(group = variable) %>%
  mutate(variable  = zissou_fit$zissou_data$x_seen_non_nested %>% colnames())

seen_year_species_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_year_species_betas")) %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())

seeing_year_species_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_year_species_betas")) %>%
  # filter(variable == "seeing_year_species_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_year_species %>% colnames())

seen_region_cluster_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seen_region_cluster_betas")) %>%
  # filter(variable == "seen_region_cluster_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())

seeing_region_cluster_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "seeing_region_cluster_betas")) %>%
  # filter(variable == "seeing_region_cluster_betas") %>%
  rename(group = variable) %>%
  mutate(variable = zissou_fit$zissou_data$x_seen_region_cluster %>% colnames())

did_betas <- zissou_fit$zissou_estimates %>%
  filter(stringr::str_detect(variable, "mpa_effect")) %>%
  # filter(variable == "mpa_effect") %>%
  mutate(group = variable) %>%
  mutate(year = zissou_fit$did_data$year %>% unique())

 betas <- bind_rows(
   seen_non_nested_betas,
   seeing_non_nested_betas,
   seen_year_species_betas,
   seeing_year_species_betas,
   seen_region_cluster_betas,
   seeing_region_cluster_betas,
   did_betas %>% select(-year)
 ) %>%
   as_data_frame()

 non_nested_beta_plot <- betas %>%
   filter(str_detect(group, "non_nested")) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_pointrange(aes(x = variable,
                       y = estimate,
                       ymin = lower,
                       ymax = upper)) +
   facet_wrap(~group) +
   coord_flip() +
   theme(axis.text.y = element_text(size = 10))

 year_species_effects_plots <- betas %>%
   filter(str_detect(group, "year_species_betas")) %>%
   mutate(year = str_replace_all(variable,"\\D","") %>% as.numeric()) %>%
   mutate(classcode = str_split(variable,'-', simplify = T)[,2]) %>%
   ggplot() +
   geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
   geom_ribbon(aes(x = year, ymin = lower, ymax = upper, fill = classcode), alpha = 0.25) +
   geom_line(aes(x = year, y = estimate, color = classcode)) +
   facet_wrap(~group)

  region_cluster_plots <- betas %>%
   filter(str_detect(group, "region_cluster_betas")) %>%
   mutate(cluster = str_replace_all(variable,"\\D","")) %>%
   mutate(region = str_split(variable,'-', simplify = T)[,3]) %>%
   ggplot() +
    geom_hline(aes(yintercept = 0), linetype = 2, color = "red") +
    geom_pointrange(aes(x = region,
                        y = estimate,
                        ymin = lower,
                        ymax = upper, color = cluster)) +
    facet_wrap(~group)

top_species <- base_run$data[[1]]$classcode %>% unique()

life_summaries <- life_history_data %>% 
  select(classcode, tm, vbgf.k, vbgf.linf) %>% 
  rename(age_mature = tm, vbk = vbgf.k, linf = vbgf.linf) %>% 
  unique() %>% 
  filter(classcode %in% top_species) %>% 
  select(-classcode) %>% 
  gather(variable,value) %>% 
  group_by(variable) %>% 
  summarise(min_val = min(value, na.rm = TRUE),
         max_val = max(value, na.rm = TRUE))

 sim_life_history <- sim_grid %>%
    mutate(
      age_mature = map_dbl(fish, "age_mature"),
      vbk = map_dbl(fish, "vbk"),
      linf = map_dbl(fish, "linf")
    ) %>%
    select(scientific_name, age_mature, vbk, linf) %>%
    unique() %>% 
   filter(age_mature > life_summaries$min_val[life_summaries$variable == "age_mature"],
          age_mature < life_summaries$max_val[life_summaries$variable == "age_mature"],
          vbk > life_summaries$min_val[life_summaries$variable == "vbk"],
          vbk < life_summaries$max_val[life_summaries$variable == "vbk"],
          linf > life_summaries$min_val[life_summaries$variable == "linf"],
          linf < life_summaries$max_val[life_summaries$variable == "linf"])

  sim_plot <- outcomes %>%
    mutate(rsize = plyr::round_any(mpa_size, 0.25),
           reff = plyr::round_any(mpa_effect,0.2)) %>%
    filter(rsize == 0.25, depletion < 0.8, depletion > 0.2,
           scientific_name %in% sim_life_history$scientific_name
           ) %>%
    mutate(year = years_protected + 2001) %>%
    filter(year %in% did_betas$year, years_protected >1) %>% 
    group_by(year, reff) %>% 
    count() %>% 
    group_by(year) %>% 
    mutate(sc = n / sum(n))

  did_plot <- did_betas %>%
    ggplot() +
    geom_tile(data = sim_plot, aes(year, reff, fill = sc), alpha = 0.75) +
    geom_vline(aes(xintercept = 2003), color = 'red', linetype = 2, size = 2) +
    geom_hline(aes(yintercept = 0)) +
    geom_pointrange(aes(
      year,
      y = exp(estimate) - 1,
      ymin = exp(lower) - 1,
      ymax = exp(upper) - 1
    ),
    size = 1.5
    ) +
    labs(x = "Year", y = "Targeted Trend Divergence") +
    scale_fill_viridis(guide = wide_hgc,
                       name = "% of Annual Sims",
                       labels = scales::percent_format(accuracy = 1),
                       option = "magma") + 
    theme(legend.position = "top",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
        )

cip_data <- pisco_data %>%
  left_join(site_data, by = c("site","side")) %>%
  filter(is.na(eventual_mpa) == F) %>%
  filter(region %in% c("ANA", "SCI","SRI",'SMI'),
         classcode %in% top_species)

# process outcomes

```

@@

<!-- \AddToShipoutPictureFG{ -->
<!--   \AtPageCenter{% or \AtTextCenter -->
<!--     \makebox[0pt]{\rotatebox[origin=c]{45}{% -->
<!--       \scalebox{5}{\texttransparent{0.3}{DRAFT}}% -->
<!--     }} -->
<!--   } -->
<!-- } -->

No-take Marine Protected Areas (MPAs), spatial regions of the ocean in which fishing is prohibited, have a long history in the management of marine resources. Traditional cultures in Oceania utilized - often temporary - MPAs as "fish banks" for times of need [@johannes1978]. Modern MPAs were first established primarily as marine analogs to the terrestrial protection of iconic landscapes like Yellowstone or Kruger National Parks [@iucn1976;@watson2014b]. Over time our goals and expectations for MPAs have evolved; while all MPAs are expected to deliver conservation benefits within their borders, many modern MPAs are also established to  bolster fish populations throughout the region in which they are located (what we term “regional-scale effects”) [@gaines2010].

Indeed, many recent agreements to expand MPAs (most proximately, the Convention on Biological Diversity’s Strategic Plan for Biodiversity, which calls for 10% of coastal waters to be protected inside MPAs by 2020 and the International Union for Conservation of Nature call for 30% by 2030) rely on the belief that well-designed MPAs will achieve benefits both within, and outside, their borders. Despite these assumptions, our collective scientific understanding of the regional-scale conservation and fishery impacts of current and future MPAs is surprisingly limited.

What is the scientific evidence of the conservation effects of MPAs? Numerous studies provide evidence that well-enforced and appropriately sized MPAs can produce conservation benefits within their borders [@lester2009;@halpern2003; @edgar2014; @gerber2005]. As these conservation benefits accrue inside MPAs, theory holds that MPAs can affect the waters beyond their borders through the spillover of adult and larval fish from the protected to the fished areas, as well as through displacement of fishing effort. Several studies have documented empirical evidence for the existence of adult or larval spillover affecting both abundance and fisheries [e.g. @goni2010; @halpern2009;@kay2012; @stobart2009; @mcclanahan2000; @russ1996; @thompson2017;@baetscher2019;@pelc2009], as well as alteration of fishing effort in reaction to [@costa2013;@mason2012;@murawski2005] and in anticipation of [@mcdermott2019] MPA placement. The potentially more important question, however, is not whether spillover occurs (it must to some degree in any realistic scenario), but what the net effects of spillover are and whether those effects are empirically detectable. From a fishery perspective, are spillover benefits sufficient to offset losses in fishing grounds and changes in responses of displaced fishers caused by an MPA? From a conservation perspective, how much does the buildup of fish inside an MPA increase biomass outside the protected area? Overall, what are the regional effects of MPAs?

@@

What Are the Regional Effects of MPAs? {.unnumbered}
----------------------

<!-- **key thing: a lot more ways for small things to happen than big things ** -->

<!-- xx need to add in discussion of other outcomes not covered here, e.g. resiliency, tourism, etc. xx -->

Much of the literature on the effects of Marine Protected Areas focuses on assessing the conservation effects within the borders of protected areas [e.g. @edgar2014]. While these within-MPA effects are vitally important for protecting rare species, biodiversity, critical habitats, and often tourism, they paint an incomplete picture of the overall population effects of MPAs. The organisms within the borders of protected areas are generally part of a broader population, the biological stock, connected through adult movement and larval dispersal. If the goal of conservationists or natural resource managers is to increase the total abundance or productivity of a resource, a broader question we should ask of MPAs is not just are there more fish inside their borders, but also how have the reserves affected abundances throughout the region in which they are located? This logic, that MPAs will have conservation benefits for most species beyond their borders, is implicit in all multilateral calls for MPA expansion.

We define the regional conservation effects of MPAs as the change in total biomass of fish (summing inside and outside of MPAs) relative to the total biomass of fish that would have occurred without the MPAs (acknowledging that other outcomes such as increased biodiversity or resiliency are also important to conservation but are beyond the scope of this analysis). Note that this definition of regional effects is in line with the recommendations of @ferraro2018, but differs from effects measured by for example differences in outcomes before and after MPAs.

@@

<!-- Numerous works have used simulation modeling to explore the effects of MPAs on outcomes such as fisheries yields and stock abundance (see @fulton2014 and references within),but these prior studies have either focused on the effect of a select set of drivers (such as dispersal rates) on outcomes, or have been highly localized assessments that incorporate multiple drivers but are tuned to one specific region. Our model captures critical bio-economic features, while still being flexible and fast enough to be generalized to a broad array of scenarios. We use this model to explore how different drivers interact to affect the conservation and fishery outcomes of MPAs. -->

### Theoretical Conservation Effects {#conservation .unnumbered}

As a thought experiment, imagine a region that has driven its fish populations to near extinction and then places 100\% of its waters inside a no-take MPA.  In that setting, we would expect the regional-scale conservation benefits to be massive. On the other hand, imagine implementing an MPA in an area where a sedentary species has been only lightly fished.  In that setting, it is possible to create a small net conservation loss if the concentration of fishing pressure outside the reserve has a greater effect than the biomass buildup inside the reserve. Within these broad bounds, numerous factors can act to affect the regional effects of MPAs. These include the scale of adult and larval dispersal relative to the size of the MPAs  [@gaines2003; @botsford2008; @difranco2018;@mcgilliard2008], the strength and timing of density dependence in the population (e.g. pre- or post-settlement), how overfished the population would be without the MPA, and how fishing activity and management responds to the implementation of the MPAs [@hilborn2004; @hilborn1992; @walters2004; @hastings2003;@gerber2003; @gerber2005; @hilborn2004a; @gaines2010; @botsford2003]. In addition, even for the same total area of MPAs, the location and spacing of the MPAs can have a profound influence on their cumulative impact through habitat and network effects [@gaines2010; @costello2010].

```{r}
short_term <- outcomes %>%
  group_by(experiment) %>% 
  filter(years_protected == max(years_protected), years_protected >= 0) %>% 
  ungroup() %>% 
  filter(!is.na(mpa_effect))

median_effect <- median(short_term$mpa_effect[short_term$year == max(short_term$year)], na.rm = TRUE)

median_pop_effect <- median(short_term$pop_effect[short_term$year == max(short_term$year)], na.rm = TRUE)

# median_fish_effect <- median(fishery_short_term$mpa_effect[short_term$year == max(short_term$year)], na.rm = TRUE)
# 
# median_fish_effect <- median(fishery_short_term$fishery_effect[short_term$year == max(short_term$year)], na.rm = TRUE)

pos_runs <- round(mean(short_term$pop_effect[short_term$year == max(short_term$year)] > 0),2)

# pos_fish_runs <- round(mean(fishery_short_term$msy_effect[fishery_short_term$year == max(fishery_short_term$year)] > 0, na.rm = TRUE),2)

pos_fish_effect <- round(mean(short_term$f[short_term$year == max(short_term$year)] > 0),2)

small_effect <- round( 100* median(short_term$mpa_effect[short_term$depletion <= 0.4]))

min_small_effect <- round( 100* min(short_term$mpa_effect[ short_term$depletion <= 0.4]))

max_small_effect <- round( 100* max(short_term$mpa_effect[ short_term$depletion <= 0.4]))

medium_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))

min_medium_effect <- round(100*min(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))

max_medium_effect <- round(100*max(short_term$mpa_effect[short_term$depletion > 0.4 & short_term$depletion <= 0.6]))

big_effect <- round(100*median(short_term$mpa_effect[short_term$depletion > 0.6]))

min_big_effect <- round(100*min(short_term$mpa_effect[short_term$depletion > 0.6]))

max_big_effect <- round(100*max(short_term$mpa_effect[short_term$depletion > 0.6]))

```

>>>>

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{eso-pic,graphicx,transparent}

\templatetype{pnasresearcharticle}  % Choose template

\title{The Regional Effects of Marine Protected Areas}

\author[a,1]{Daniel Ovando}
\author[b]{Jennifer E. Caselle}
\author[b]{Christopher Costello}
\author[b]{Olivier Deschenes}
\author[b]{Steven D. Gaines}
\author[a]{Ray Hilborn}
\author[b]{Owen Liu}

  \affil[a]{University of Washington, School of Aquatic and Fishery Science}
  \affil[b]{University of California, Santa Barbara}

% Please give the surname of the lead author for the running footer
\leadauthor{Ovando}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Healthy marine ecosystems are critical to the well-being of the planet.
Marine protected areas, parts of the oceans protected from human
activities such as fishing, are increasingly being used in an effort to
conserve and manage these ecosystems. Our study pairs theory with
empirical methods to examine the regional effects of MPAs on
conservation and fishery outcomes. Using simulations, we show that the
effects of marine protected areas can be highly variable and dependent
on human drivers. Building off of MPA theory, we find no statistically
clear effect of a network of MPAs, which our simulation analysis
suggests is entirely reasonable. Our results lay a foundation for future
research on the design and monitoring of marine protected areas.}

\authorcontributions{D.O., S.G., and R.H. developed structure of simulation model. D.O.,
O.D., C.C., and J.C. developed estimation strategy. J.C. provided
support on collecting and interpreting data. All simulations,
statistics, and sensitivity analyses performed by D.O. and O.L. . All
authors contributed to the manuscript.}

\authordeclaration{R.H's research program receives funding from environmental NGOs,
foundations, fishing industry, governments and international agencies.
All of these can be interpreted as a conflict of interest when
evaluating fisheries policy. C.C., S.G., and O.D's research program
includes funding from environmental NGOs and foundations with an
interest in ocean conservation and food security}

\correspondingauthor{\textsuperscript{} }

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{  Marine Protected Areas |  Conservation |  Bio-economic modeling |  Program Evaluation |  Channel Islands National Marine Sanctuary  }

\begin{abstract}
Marine Protected Areas (MPAs) cover between 3-7\% of the world's oceans,
up from less than 1\% in the year 2000. The Convention on Biological
Diversity calls for 10\% of coastal waters to be protected inside MPAs
by 2020, while the International Union for Conservation of Nature calls
for 30\% protection by 2030. It is often clear that MPAs produce
conservation benefits inside their borders, but many MPAs are also
justified on the grounds that they also benefit the broader region
outside their borders. Despite the expanding role of MPAs in marine
resource management, we lack a clear understanding of what the regional
effects of MPAs might be, and how we might detect them in scientific
studies. We used a bio-economic simulation model to demonstrate how the
regional effects of MPAs are affected by a suite of economic and
environmental drivers and show how commonly used methods to measure
inside-MPA effects are likely to fail to capture the regional impacts of
protection. We present an empirical strategy for estimating regional MPA
effects on biomass density using 16 years of data from inside and
outside of MPAs within the Channel Islands, California, USA. We find no
statistically clear effect of MPAs on aggregate mean biomass densities
of targeted finfish throughout the study region. To make meaningful
progress on MPA design and evaluation, the next generation of MPA
science must address the challenge of understanding and estimating the
regional effects of MPAs across the diversity of species in the
ecosystem.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\acknow{This study would not be possible without the work provided by PISCO
divers over the years. DO thanks Cody Szuwalski, Julia Lawson, and André
Punt for helpful comments and technical support.}

@@

No-take Marine Protected Areas (MPAs), spatial regions of the ocean in
which fishing is prohibited, have a long history in the management of
marine resources. Traditional cultures in Oceania utilized - often
temporary - MPAs as ``fish banks'' for times of need (1). Modern MPAs
were first established primarily as marine analogs to the terrestrial
protection of iconic landscapes like Yellowstone or Kruger National
Parks (2, 3). Over time our goals and expectations for MPAs have
evolved; while all MPAs are expected to deliver conservation benefits
within their borders, many modern MPAs are also established to bolster
fish populations throughout the region in which they are located (what
we term ``regional-scale effects'') (4).

Indeed, many recent agreements to expand MPAs (most proximately, the
Convention on Biological Diversity's Strategic Plan for Biodiversity,
which calls for 10\% of coastal waters to be protected inside MPAs by
2020 and the International Union for Conservation of Nature call for
30\% by 2030) rely on the belief that well-designed MPAs will achieve
benefits both within, and outside, their borders. Despite these
assumptions, our collective scientific understanding of the
regional-scale conservation and fishery impacts of current and future
MPAs is surprisingly limited.

What is the scientific evidence of the conservation effects of MPAs?
Numerous studies provide evidence that well-enforced and appropriately
sized MPAs can produce conservation benefits within their borders
(5--8). As these conservation benefits accrue inside MPAs, theory holds
that MPAs can affect the waters beyond their borders through the
spillover of adult and larval fish from the protected to the fished
areas, as well as through displacement of fishing effort. Several
studies have documented empirical evidence for the existence of adult or
larval spillover affecting both abundance and fisheries (9--17), as well
as alteration of fishing effort in reaction to (18--20) and in
anticipation of (21) MPA placement. The potentially more important
question, however, is not whether spillover occurs (it must to some
degree in any realistic scenario), but what the net effects of spillover
are and whether those effects are empirically detectable. From a fishery
perspective, are spillover benefits sufficient to offset losses in
fishing grounds and changes in responses of displaced fishers caused by
an MPA? From a conservation perspective, how much does the buildup of
fish inside an MPA increase biomass outside the protected area? Overall,
what are the regional effects of MPAs?

@@

\hypertarget{what-are-the-regional-effects-of-mpas}{%
\subsection*{What Are the Regional Effects of
MPAs?}\label{what-are-the-regional-effects-of-mpas}}
\addcontentsline{toc}{subsection}{What Are the Regional Effects of
MPAs?}

Much of the literature on the effects of Marine Protected Areas focuses
on assessing the conservation effects within the borders of protected
areas (7). While these within-MPA effects are vitally important for
protecting rare species, biodiversity, critical habitats, and often
tourism, they paint an incomplete picture of the overall population
effects of MPAs. The organisms within the borders of protected areas are
generally part of a broader population, the biological stock, connected
through adult movement and larval dispersal. If the goal of
conservationists or natural resource managers is to increase the total
abundance or productivity of a resource, a broader question we should
ask of MPAs is not just are there more fish inside their borders, but
also how have the reserves affected abundances throughout the region in
which they are located? This logic, that MPAs will have conservation
benefits for most species beyond their borders, is implicit in all
multilateral calls for MPA expansion.

We define the regional conservation effects of MPAs as the change in
total biomass of fish (summing inside and outside of MPAs) relative to
the total biomass of fish that would have occurred without the MPAs
(acknowledging that other outcomes such as increased biodiversity or
resiliency are also important to conservation but are beyond the scope
of this analysis). Note that this definition of regional effects is in
line with the recommendations of (22), but differs from effects measured
by for example differences in outcomes before and after MPAs.

@@

\hypertarget{conservation}{%
\subsubsection*{Theoretical Conservation Effects}\label{conservation}}
\addcontentsline{toc}{subsubsection}{Theoretical Conservation Effects}

As a thought experiment, imagine a region that has driven its fish
populations to near extinction and then places 100\% of its waters
inside a no-take MPA. In that setting, we would expect the
regional-scale conservation benefits to be massive. On the other hand,
imagine implementing an MPA in an area where a sedentary species has
been only lightly fished. In that setting, it is possible to create a
small net conservation loss if the concentration of fishing pressure
outside the reserve has a greater effect than the biomass buildup inside
the reserve. Within these broad bounds, numerous factors can act to
affect the regional effects of MPAs. These include the scale of adult
and larval dispersal relative to the size of the MPAs ({\textbf{???}},
23--25), the strength and timing of density dependence in the population
(e.g.~pre- or post-settlement), how overfished the population would be
without the MPA, and how fishing activity and management responds to the
implementation of the MPAs (4, 8, 26--32). In addition, even for the
same total area of MPAs, the location and spacing of the MPAs can have a
profound influence on their cumulative impact through habitat and
network effects (4, 33).

We tested the effects of different combinations of these theoretical
drivers of MPA conservation effects using over 10,000 simulated MPA
scenarios. A striking outcome of this simulation analysis is how the
level of fishing pressure can cause dramatic and predictable shifts in
conservation outcomes. We measure fishing pressure by the degree of
simulated depletion (percent of unfished biomass gone, such that a
depletion of 0\% means the population is at 100\% of unfished biomass,
100\% means 0\% of unfished biomass remains) in the scenario without
MPAs. When MPAs were placed in relatively unexploited fisheries
(depletions less than 40\%), the median regional conservation effect was
barely noticeable (4\%). For ecosystems fished near MSY levels
(depletion of 40\%-60\%), the median conservation effect was 28\%, while
for overfished ecosystems (without-MPA depletions above 60\%) the median
regional conservation effect was over 100\%
(Fig.\ref{conservation-effects}-A). Across all of these groups though,
effects as low as declines of -50\% and as high as an increase of more
than 100\% were possible. It should be noted though that these runs
represent the range of scenarios evaluated in our simulation, and that
certain types of scenarios may be much more likely than others in the
real world. To put these results into context, the FAO estimates that
7\% of the worlds fisheries with status estimates fall into the
relatively unexploited category, 60\% fall into the fully fished
category, and 33\% fall into the heavily fished category (34), though
works that include a broader range of fisheries estimate that 50\% or
more of stocks to fall into the heavily fished category (35, 36).

These simple results support the intuitive conclusion that bigger MPAs
on overfished populations produce large positive conservation outcomes.
However, when combined with other factors such as fleet dynamics, life
history, and MPA size, the conservation effects of MPAs can vary widely
in magnitude and timing, a finding corroborated by empirical evidence
(37--40) (Fig.\ref{conservation-effects}-B). MPAs protecting small
proportions of a stock can produce small positive effects, large
positive effects, or even small negative effects. As the area protected
inside MPAs increases, positive effects become more likely, but even
MPAs protecting most of a population's footprint can produce
conservation effects as low as 0\% and as high as 200\% and above
depending on the context in which they are placed. The degree of
depletion (i.e.~fishing pressure) has a clearer signal, with small
conservation benefits of MPAs when depletion is low, and larger effects
when depletion is high, but again even for severely depleted populations
the conservation effects of MPAs can vary widely (although nearly always
with population benefits).
